---
title: "Final Project: Covid 19 Around the World"
description: |
  "By using owid Covid 19 data and statistics, I examined trends."
site: distill::distill_website
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(tidyverse)
library(plotly)
```

```{r}
raw_data <- read_csv(file="owid-covid-data.csv")
```

```{r Covid-plot}

clean_data <-raw_data %>% 
  select(continent, total_cases) %>% 
  drop_na() %>% 
  group_by(continent) %>% 
  summarize(continent_cases=sum(total_cases)) 
clean_data %>% 
  ggplot(mapping=aes(x=continent,y=continent_cases,fill=continent))+
  geom_col()+
  scale_y_continuous(labels = scales::number_format())+
  labs(title="Covid Cases per Continent", x="Continent",y="Continent Cases",fill="Continents")

```

```{r}
clean_data1 <- raw_data %>% 
  select(continent, total_deaths) %>% 
  drop_na() %>% 
  group_by(continent) %>% 
  summarize(continent_deaths=sum(total_deaths)) 
clean_data1 %>% 
  ggplot(mapping=aes(x=continent,y=continent_deaths,fill=continent))+
  geom_col()+
  scale_y_continuous(labels = scales::number_format())+
  labs(title="Covid Deaths per Continent", x="Continent",y="Continent Deaths",fill="Continents")
```

-There is no data on the continent of Antarctica because of insufficient human population 

-Furthermore, there is very little cases in Oceania/Australia that I will consider it to be neglible for data and filter out in the following graph

-Initial Data shows a shockingly low amount of Covid Cases/Deaths in Africa compared to other continents


```{r}
clean_data2 <- raw_data%>% 
  filter(continent==c("Africa","Asia","Europe","North America","South America")) %>% 
  select(date, new_cases, continent, location) %>% 
  drop_na(new_cases) %>% 
  group_by(continent, date) %>% 
  summarize(new_case=sum(new_cases)) %>% 
  ggplot(aes(x=date, y=new_case,color=continent, text=paste("Date:", date,"\n","New Case:",new_case)))+
  geom_point()+
  scale_y_continuous(labels = scales::number_format())+
  labs(title="Weekly New Cases Among the Continents Throughout the\nYear", x="Date", y="New Cases", color="Continent", caption="Source: Our World in Data")

ggplotly(clean_data2,tooltip="text")

```
<aside>
Source: Our World in Data
</aside>

The plot above has all the continents in one graph allowing for easier comparison. The plot above separates the continents for a cleaner look. Also: The plots are interactive, hovering over points will provide precise dates and new cases.

```{r}
clean_data2 <- raw_data%>% 
  filter(continent==c("Africa","Asia","Europe","North America","South America")) %>% 
  select(date, new_cases, continent, location) %>% 
  drop_na(new_cases) %>% 
  group_by(continent, date) %>% 
  summarize(new_case=sum(new_cases)) %>% 
  ggplot(aes(x=date, y=new_case,color=continent, text=paste("Date:", date,"\n","New Case:",new_case)))+
  geom_point()+
  facet_wrap(~continent)+
  scale_y_continuous(labels = scales::number_format())+
  labs(title="Weekly New Cases Among the Continents Throughout the\nYear",x="Date", y="New Cases", color="Continent", caption="Source: Our World in Data")

ggplotly(clean_data2,tooltip="text")

```
<aside>
Source: Our World in Data
</aside>

-North America and Asia has surges in new cases in January 2021, and May 2021 respectively

-Both North America and Asia made significant progress in decreasing the new cases

-Africa, South America, and Europe have been relatively consistent in new cases 

-Europe is progressing in decreasing new cases, but South America has seen an increase recently


